---
title: "Predict on the Weight Lifting Exercise"
author: "Chloe Kim"
date: '2020 3 24 '
output: html_document
---
## Summary
Purpose of this study is constructing prediction model to classify 5 different ways of dumbell lifting and 20 cases in the test data set would be tested by the prediction model. For this classification modeling, randomForest was applied and data from accelerometers,gyroscope and magnetometer on the belt,forearm, arm  and dumbell were selected as feature.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Data
Download training data from link and explore data set with additional information from the website. (http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har)
Classification outcome is "classe" variable with 5 levels, A to E. 

```{r}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",destfile ="pml-training.csv")
training <- read.csv("pml-training.csv",na.strings = "na")
summary(training$classe)
```

## Feature selection
Data of accelerometer, gyroscope and magnetometer readings were selected for modeling. 
According to the published WLE dataset paper (Velloso et al. ACM SIGCHI,2013) roll, pitch, yaw of belt were seletced additionally as feature.  
Total 35 features (11 in belt, 8 in arm, dumbbell and forearm repectively) were selected for RandomForest training. 

```{r}
suppressMessages(library(caret))
suppressMessages(library(randomForest))
suppressMessages(library(dplyr))
## Feature select and correct variable type
  features <- grep("accel_|gyros_|magent_",names(training)) 
  dt <- training[,c(8:10,features,160)]
  nms <- c("var_total_accel_belt","var_accel_arm","var_accel_dumbbell","var_accel_forearm")
  dt[nms] <- lapply(dt[nms],as.numeric)
## Subset into train and test set, 0.6 to 0.4 ratio  
  set.seed(338) 
  trainid <- createDataPartition(dt$classe, p=0.6, list=F) 
  train <- dt[trainid,] ## 11776 obs. of 36 variables
  test <- dt[-trainid,] ## 7846 obs. of 36 variables
## RandomForest modeling 
  set.seed(1101)
  rf <- randomForest(classe~., data=train)
## Extract the variable importance and oder
  varImp <- data.frame(var=row.names(varImp(rf)), overall=varImp(rf)) %>% arrange(desc(Overall))
  tail(varImp,10)
```

Variances of accelerometer(var_total_accel_belt,var_accel_arm,var_accel_dumbbell,var_accel_forearm) could be excluded according to the Importance measure but they were retained at this time.

### Performance : Prediction and Accuracy
In rf model, Out of bag error was estimate 1.09% and addtional cross validtion was not performed due to intrinsic basis of randorm forest analysis. 
```{r}
## OBB error rate and confusion matrix of model
print(rf)
 
```
 
  From the confusionMatrix of internal test set with rf model, overall accuracy is observed 98.9%.
```{r}
## Test model accuraty with internal test set 
  rfprd <- predict(rf,test[-36])
  cM <- confusionMatrix(test$classe,rfprd)
  cM$overall
  cM$table
  
```


### Quiz : apply prediction model to 20 test cases in the test data. 
Using this rf model, classes of 20 cases in the test data set were prediected. 

```{r}
## Load the test data set and subset data frame with selected features  
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",destfile ="pml-testing.csv")
  finaltest <- read.csv("pml-testing.csv",na.strings = "na")
  features <- grep("accel_|gyros_|magent_",names(finaltest)) 
  data <- finaltest[,c(8:10,features,160)]
  nms <- c("var_total_accel_belt","var_accel_arm","var_accel_dumbbell","var_accel_forearm")
  data[nms] <- lapply(data[nms],as.numeric)
  
## Prediction of 20 casese with RandomFores model rf.
  rfprdfinal <- predict(rf,data[-36]) 
  print(rfprdfinal)
```
